---
title: "COVID-19 in Mexico: Visualizing the municipalities of hope"
output: html_notebook
---

By summer 2020, the federal government in Mexico presented a program that would lead the re-opening of the economy, *_"municipios de la esperanza"_* (municipalities of hope). This notebook shows the execution process of building a visualization with which better evaluate the extent of this  program in Mexico. There are two main parts to this process, data transformation and cartogram plotting. 

## Data transformation

```{r}
# Loading packages
library(stringr)
library(dplyr)
```



```{r}
mun_esp <- read.csv("municipios_esperanza.csv")

```

## Cartogram plotting

```{r}
library(sf)
library(cartogram)
library(dplyr)
library(tmap)
library(tmaptools)
library(ggplot2)
library(viridis)

wd <- "C:/Users/santa/Desktop/COVID"
setwd(wd)

muni_data <- read.csv("municipios_esperanza.csv")
map_data <- st_read("mun2019gw.shp", stringsAsFactors = FALSE)

map_data$CVEGEO <- as.numeric(map_data$CVEGEO)
muni_data$POB_TOT_2020 <- as.numeric(muni_data$POB_TOT_2020)

combo <- inner_join(map_data, muni_data, by = "CVEGEO")
combo <- st_transform(combo, 6372)

mapa <- cartogram_cont(combo, "POB_TOT_2020", itermax = 20)
mapa_ncont <- cartogram_ncont(combo, "POB_TOT_2020", k = 20)

tm_shape(combo) +
  tm_fill("esp", palette = "-plasma") +
  tm_layout(frame = FALSE, legend.position = c("left", "bottom"))

tm_shape(map_data) + tm_borders(alpha = 0.03) +
  tm_shape(mapa_ncont) + tm_polygons(lwd = NA, "esp", palette = "-viridis") +
  tm_layout(frame = FALSE, legend.show = FALSE)
```

